---
title: "swissLipidsR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{swissLipidsR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# How to use <code>SwissLipidsR</code>

SwissLipids is one of the major databases for lipids and lipid related projects.
It offers an API for querying the database. <code>SwissLipidsR</code> is a package
to interact with this API. It is based on <code>jsonlite</code> to interact with
the webservice. This vignette explains the basic functions.

```{r setup, message=FALSE}
library(SwissLipidsR)
```

## Mapping of identifiers

The SwissLipids API offers to map between different lipid related database identifier.
Identifier from Swisslipids, LipidMaps, ChEBI or HMDB can be mapped to identifer
from SwissLipids, LipidMaps, ChEBI, HMDB, UniProtKB or Rhea. The following example
maps from SwissLipid IDs to LipidMaps IDs.

```{r}
# define mapping parameters
ids <- c("SLM:000048885", "SLM:000000651")
from <- "SwissLipids"
to <- "LipidMaps"

# perform mapping
results <- swissLipidsMapping(from, to, ids)

results
```

The results data frame contains a second nested data frame. This can be unnested
using functions from the <code>tidyverse</code> package.

```{r tidyverse, message=FALSE}
library(tidyverse)
```

The function <code>unnest()</code> performs the required operation.

```{r}
unnest(results, cols = c(to))
```

## Perform searches

Beside identifier mapping the SwissLipids API offers different searching functionalities.
The basic search can be used to get general information on possible entries.

```{r}
results <- swissLipidsSearch("Phosphatidate (36:2)")
```

Different advanced search options are available, e.g. searching by name, SMILES,
InChIKey, formula or mass. The first example searches all results that match the
lipid shorthand notation PC(34:2).

```{r}
results <- swissLipidsAdvancedSearch(name = "PC(34:2)")

head(results)
```

The second example performs a sum formula search. This search might return results
from different levels of the lipid classification. The additional function
<code>filterResults</code> can be used to filter the results based on the 
classficiation level.

```{r}
results <- swissLipidsAdvancedSearch(formula = "C39H76NO8P")

head(results)

results <- swissLipidsAdvancedSearch(formula = "C39H76NO8P") %>% 
  filterResults(classification_level = "Species")

head(results)
```

Also searches on masses can be performed. The search mass can be either a neutral
mass ("MassExact") or an adduct mass. Currently supported adducts are [M]+ ("MassM"),
[M+H]+ ("MassMH"), [M+K]+ ("MassMK"), [M+Na]+ ("MassMNa"), [M+Li]+ ("MassMLi"),
[M+NH4]+ ("MassMNH4"), [M-H]- ("MassMmH"), [M+Cl]- ("MassMCl"), [M+HAc-H]- ("MassMOAc").

```{r}
results <- swissLipidsAdvancedSearch(mz = 410.243,
                                     adduct = "MassExact",
                                     massErrorRate = 0.001)

head(results)
```

